---
output:
  word_document: default
  pdf_document: default
  html_document: default
---

## Capstone Project Proposal
###### *Introduction to Data Science by Springboard*
### Network Data Flow analysis
##### By Gabriel Fontenot
***

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
```

### Cleaning

# 0: Reference required Libraries in RStudio
```{r library}

library(dplyr)
library(tidyr)
library(stringr)
library("rjson")
```

# 1: Load the 'flow' source data set into RStudio

```{r load_flows}

dat = read.csv("smallFlows_clean.csv", header = TRUE)
```

# 2: Counting

Count the number of rows in the data set.
```{r count_rows}
count <- nrow(dat)
print(count)
```

Count the number of rows in the data set that have a platform of "CentOS-6.5"

```{r count_rows_platform}

count_filter <- nrow(filter(dat, platform == "CentOS-6.5"))
print(count_filter)
```

# 3: Trends (high, low, increase, decrease, anomalies)

```{r trends_anomalies}

a <- dat %>% group_by(platform) %>% tally
ggplot(a, aes(x=platform, y=n,)) + 
  geom_point(na.rm=TRUE, aes(colour = platform))
```

# 4: Bar plot or a histogram

```{r bar_plot_histogram}

ggplot(dat,aes(x=Length, group=platform,fill=platform)) +
  geom_histogram(na.rm=TRUE, position="dodge", binwidth=500) + 
  theme_bw()
```

# 5: Compaing Quantities

```{r compare_quantities}

ggplot(dat, aes(x=Time, colour=Protocol, fill=Protocol)) + 
  geom_density(alpha = 0.1) +
  xlim(235, 240)
```

# 6: Scatterplot

```{r scatterplot}

ggplot(dat, aes(x = Length, y = platform, colour = cond)) + 
  geom_point(na.rm=TRUE, aes(colour = platform)) + 
  theme(legend.position="none")
```

# 7: Time Series Plot

```{r time_series_plot}

# created_at
# last_config_fetch_at

ggplot(dat, aes(x=as.POSIXct(created_at, origin="1970-01-01"), y=platform,)) + 
  geom_point(na.rm=TRUE, aes(colour = platform)) + 
  theme(legend.position="none")

number_ticks <- function(n) {function(limits) pretty(limits, n)}

ggplot(dat, aes(x=Time, y=platform)) + 
  geom_point(na.rm=TRUE, aes(colour = platform)) +
  scale_x_continuous(breaks=number_ticks(20)) + 
  theme(legend.position="none")
```

# 8: MISC

```{r misc}

ggplot(dat, aes(x=Time, y=platform, color=platform)) +
  geom_boxplot(na.rm=TRUE)
```

# 9: Submit the project on Github
