---
output:
  word_document: default
  pdf_document: default
  html_document: default
---

## Capstone Project Proposal
###### *Introduction to Data Science by Springboard*
### Network Data Flow analysis
##### By Gabriel Fontenot
***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Cleaning

# 0: Reference required Libraries in RStudio
```{r library}
library(dplyr)
library(tidyr)
library(stringr)
library("rjson")
```

# 1: Load the 'flow' source data set into RStudio

```{r load_flows}
tst = read.csv("smallFlows.csv", header = TRUE)
```

# 2: Remove ROWs with non IPv4 addresses in Source and Destination

```{r remove_rows}
tst <- tst %>% 
  filter(str_detect(Source, ".*?(\\d+\\.\\d+\\.\\d+\\.\\d+).*?")) %>%
  filter(str_detect(Destination, ".*?(\\d+\\.\\d+\\.\\d+\\.\\d+).*?"))
```

# 3: Load the 'host' source data set into RStudio

```{r load_hosts}
hostName = read.csv("hostName.csv", header = TRUE)
```

# 4: Merge the flows and host data sets

```{r merge_flows_hosts}
tst <- tst %>% 
  right_join(hostName, c("Source" = "ip"))
```

# 5: View and save results

```{r view_results}
print(tst)
print(hostName)
```

```{r write_results}
write.csv(tst, file = "smallFlows_clean.csv")
```

# 6: Submit the project on Github
