---
title: "Machine Learning"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(rjson)

```

## Preparing the Dat Sets

The following section prepares the two data sets that will be used for the machine learning exercise. The data provided must be cleaned and prepared to allow for proper analysis.

```{r data_preparation}
# Loading the data sets

fileName <- "/Users/gfonteno/data/Rstudio/springboard/0.capstone-project/FlowData.csv"
flowData = read.csv(fileName, header = TRUE)

fileName <- "/Users/gfonteno/data/Rstudio/springboard/0.capstone-project/SensorData.csv"
sensorData = read.csv(fileName, header = TRUE)

# Cleaning the flow/host data sets

#1) Remove ROWs with non IPv4 addresses in Source and Destination

flowData <- flowData %>% 
  filter(str_detect(src_address, ".*?(\\d+\\.\\d+\\.\\d+\\.\\d+).*?")) %>%
  filter(str_detect(dst_address, ".*?(\\d+\\.\\d+\\.\\d+\\.\\d+).*?"))

#2) Remove ROWs with no (missing) destination HOSTNAME

flowData <- with(flowData, flowData[!(dst_hostname == "" | is.na(dst_hostname)), ])

#3) Merge data sets, flows and hosts

data <- merge(sensorData, flowData, by.x=c('host_name'), by.y=c('dst_hostname'))

#4) Add new column Total Packet, total_pkts

data$total_pkts <- data$rev_pkts + data$fwd_pkts

#5) Add new column Total Bytes, total_bytes

data$total_bytes <- data$rev_bytes + data$fwd_bytes

#6) Add new column Platform Type, platform_type

data <- data %>% 
  mutate(platform_type = ifelse(str_detect(platform, ".*?(^MS.+).*?"), 0, 1))

#7) Add new column time, time_min

data <- data %>% mutate(time_min = format(as.POSIXct(strptime(data$timestamp,"%Y-%m-%dT%H:%M:%S",tz="")) ,format = "%I:%M"))

```

## Machine Learning applied to Capstone project

A logistic regression model will be used with idependent varialbe 'platform_type' to predict the probablity of that a flow will be sourced to a specific host type, Windows or Linux.

One of the first exercises in evaluating the overall probablility will involve looking into the class bias. This will help us understand the expected predicted value.

We will create a test and a training set of the data model to reduce the overall size of the evalutated data that will later be tested against the test set.

```{r pressure, echo=FALSE}
#Logistic Regrsssion

# Generate a TEST AND TRAINING SET
set.seed(100)
split = sample.split(data$platform_type, SplitRatio = 0.55)
dataTrain = subset(data, split == TRUE)
nrow(dataTrain)
dataTest = subset(data, split == FALSE)
nrow(dataTest)

# Determine the total for each field of the dependent variable
table(dataTrain$platform_type)

# Update the dependent variable field to convert it to a vector
data$platform_type <- as.factor(data$platform_type)

# Logistic Regression
platformLog = glm(platform_type ~ proto + total_pkts, data = dataTrain, family = binomial)
summary(platformLog)

# Prediction
predictTrain = predict(platformLog, type="response")
summary(predictTrain)

# Display the mean of the Training set
tapply(predictTrain, dataTrain$platform_type, mean)

# 1 is Windows Platform
# 0 is Linux
```

The output of the analysis infers that there is a hight probability that the flow will be of type Windows as compared to Linux. The analysis is based on independent variables 'proto' and 'total_pkts'.
